---
title: "Analyzing circadian rhythms with circadiandynamics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyzing circadian rhythms with circadiandynamics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(circadiandynamics)
```

## Reading data

Each format has one reader (`read_trikinetics()`, `read_clocklab()`, `read_csv_data()`, `read_satellite()`, `read_vitalpatch()`). Pass a file or a folder; leave the path `NULL` to pick one via a dialog. Trikinetics supports a nested per-individual `layout` for analysis and a `"long"` layout for actograms.

```{r, eval = FALSE}
# nested per-individual list, for analysis:
monitor <- read_trikinetics("path/to/monitor.txt")
# long form, for actograms:
actogram(read_trikinetics("path/to/monitor.txt", layout = "long")[-2], sampling = 1)
```

## A worked example on the bundled dataset

`trikinetics` is a bundled example (32 individuals, 1-min resolution). We bin to 15 minutes (plenty for circadian period estimation) and analyze two individuals.

```{r}
data(trikinetics)
library(lubridate); library(dplyr)
bin15 <- function(d) d |>
  mutate(datetime = floor_date(datetime, "15 minutes")) |>
  group_by(datetime) |>
  summarise(value = sum(value), .groups = "drop")

sub <- lapply(trikinetics[c("IND 2", "IND 4")], bin15)
result <- process_timeseries_main(
  sub, make_windows = FALSE, sampling_rate = "15 minutes",
  from = 15, to = 33, detrend_data = FALSE, butterworth = TRUE,
  f_low = 1/(12*4), f_high = 1/(35*4), order = 2, ofac = 1
)
periods <- data.frame(
  individual = names(result),
  acf_period_h = sapply(result, \(x) x$acf$results$period),
  lsp_period_h = sapply(result, \(x) x$lomb$results$period)
)
knitr::kable(periods)
```

## Tidying and multiple-testing correction

```{r}
tidy <- simplify_data(result)
names(tidy)
# Benjamini-Hochberg FDR across the Lomb-Scargle p-values:
adjust_pvalues(tidy$lombscargle, p_col = "cosinor_p_value")$cosinor_p_value_adj
```

## Plots

`detailed_plots()` writes per-individual figures, and `simplify_data()` output feeds the `plot_*` family (e.g. `plot_actogram_windows()`, `plot_lsp_results()`).

```{r, eval = FALSE}
detailed_plots(result, sampling_rate = "minutes", windows = FALSE)
```
